systemfit: A Package for Estimating\\ Systems of Simultaneous Equations in \proglang{R}

### 2006-08-30 paper/trunk/systemfit.tex
added footer for nested docs
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@632 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changes according to the first major comment of reviewer A: modified title and a few sentences (package to estimate -> package for estimating; equation system -> system of equations)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@691 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\title{systemfit: A Package to Estimate\\
       Simultaneous Equation Systems in \proglang{R}}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Arne Henningsen, Jeff D.\ Hamann} %% comma-separated
\Plaintitle{systemfit: A Package to Estimate Simultaneous Equation Systems 
            in R} %% without formatting
\Shorttitle{systemfit} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{ 
Many statistical analyses are based on models containing systems of
structurally related equations.
In cases where cross-equation disturbances are correlated,
full information methods are required \citep{zellner62}.
If exogenous variables are stochastically dependent
on the disturbances in the system,
then instrumental variable estimation methods should be used
\citep{zellner62b}
The package \pkg{systemfit} provides the capability
to estimate systems of linear equations
within the \proglang{R} programming environment.
}

\Keywords{R, simultaneous equations systems,
  seemingly unrelated regression, two-stage least squares,
  three-stage least squares, parameter restriction}

%% at least one keyword must be supplied

%% publication information
%% NOTE: This needs to filled out ONLY IF THE PAPER WAS ACCEPTED.
%% If it was not (yet) accepted, leave them commented.
%% \Volume{13}
%% \Issue{9}
%% \Month{September}
%% \Year{2004}
%% \Submitdate{2004-09-29}
%% \Acceptdate{2004-09-29}

@@

% \section{CVS Revision Log}
% \begin{verbatim}
% $Log$
% Revision 1.19  2005/11/08 06:31:48  hamannj
% made seemingly massive changes... but hopefully not for the worse...
%
% Revision 1.18  2005/11/07 14:08:56  henningsena
% a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
%
% Revision 1.17  2005/11/07 10:21:59  henningsena
% just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
%
% Revision 1.16  2005/11/02 18:31:45  hamannj
% small last changes
%
% Revision 1.15  2005/11/02 18:15:40  hamannj
% small changes

% Revision 1.14  2005/10/26 21:16:53  hamannj
% checked out and updated local copy before edits

% Revision 1.13  2005/09/26 20:34:12  henningsena
% using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)

% Revision 1.12  2005/09/26 11:24:05  henningsena
% some minor changes

% Revision 1.11  2005/09/01 12:27:55  henningsena
% revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.

% Revision 1.10  2005/09/01 11:49:02  henningsena
% moved section about non-linear estimation from systemfit.tex into new file nlsystemfit.tex, because this section should not be included in (this version of) the article

% Revision 1.9  2004/11/16 10:59:09  henningsena
% some minor changes and some comments on nlsystemfit

% Revision 1.8  2004/11/16 08:55:53  henningsena
% merged file and resolved conflict due to some minor changes I did while Jeff was adding nlsystemfit.

% Revision 1.7  2004/11/16 05:34:47  hamannj
% question about nlsystemfit code section (example)

% Revision 1.6  2004/11/16 05:32:46  hamannj
% made small changes because my code section wouldn't compile. argh.

% Revision 1.5  2004/11/16 05:31:19  hamannj
% added cvs header and log section. to be removed when submitted.

% \end{verbatim}

@@

>>>>

\title{systemfit: A Package for Estimating\\
       Systems of Simultaneous Equations in \proglang{R}}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Arne Henningsen, Jeff D.\ Hamann} %% comma-separated
\Plaintitle{systemfit: A Package for Estimating Systems of Simultaneous Equations
            in R} %% without formatting
\Shorttitle{systemfit} %% a short title (if necessary)

%% an abstract and keywords
\input{systemfit_abstract.tex}

%% publication information
%% NOTE: This needs to filled out ONLY IF THE PAPER WAS ACCEPTED.
%% If it was not (yet) accepted, leave them commented.
%% \Volume{13}
%% \Issue{9}
%% \Month{September}
%% \Year{2006}
%% \Submitdate{2006-02-28}
%% \Acceptdate{2006-09-29}

@@

@@

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2006-08-30 paper/trunk/systemfit_intro.tex
changes according to the first major comment of reviewer A: modified title and a few sentences (package to estimate -> package for estimating; equation system -> system of equations)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@691 591db2bd-e0a0-4d22-8c19-2192dbb7da31

Many theoretical models consist of more than one equation.
If these models are econometrically estimated,
the disturbance terms of these equations are likely
to be contemporaneously correlated,
because unconsidered factors
that influence the disturbance term in one equation
probably influence the disturbance terms in other equations.
Ignoring this contemporaneous correlation
and estimating these equations separately
leads to inefficient parameter estimates.
However, estimating all equations simultaneously,
taking the covariance structure of the residuals into account,
leads to efficient estimates.
This estimation procedure is generally called
°Seemingly Unrelated Regression° (SUR) \citep{zellner62}.
Another reason to estimate an equation system simultaneously are
cross-equation parameter restrictions.%
\footnote{
Especially the economic theory suggests many cross-equation parameter
restrictions (e.g.\ the symmetry restriction in demand models).
}
These restrictions can be tested and/or imposed only in a simultaneous
estimation approach.

@@

The \pkg{systemfit} package provides the capability to estimate
linear equation systems in \proglang{R}
\citep{r-project}.
Although linear equation systems can be estimated
with several other statistical and econometric software packages
(e.g.\ \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
\pkg{systemfit} has several advantages.
First, all estimation procedures are publicly available in the source code.
Second, the estimation algorithms can be easily modified to meet specific
requirements.
Third, the (advanced) user can control estimation details generally
not available in other software packages by overriding reasonable defaults.

>>>>

Many theoretical models consist of more than one equation.
If these models are econometrically estimated,
the disturbance terms of these equations are likely
to be contemporaneously correlated,
because unconsidered factors
that influence the disturbance term in one equation
probably influence the disturbance terms in other equations.
Ignoring this contemporaneous correlation
and estimating these equations separately
leads to inefficient parameter estimates.
However, estimating all equations simultaneously,
taking the covariance structure of the residuals into account,
leads to efficient estimates.
This estimation procedure is generally called
°Seemingly Unrelated Regression° (SUR) \citep{zellner62}.
Another reason to estimate a system of equations simultaneously are
cross-equation parameter restrictions.%
\footnote{
Especially the economic theory suggests many cross-equation parameter
restrictions (e.g.\ the symmetry restriction in demand models).
}
These restrictions can be tested and/or imposed only in a simultaneous
estimation approach.

@@

The \pkg{systemfit} package provides the capability to estimate
systems of linear equations in \proglang{R}
\citep{r-project}.
Although systems of linear equations can be estimated
with several other statistical and econometric software packages
(e.g.\ \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
\pkg{systemfit} has several advantages.
First, all estimation procedures are publicly available in the source code.
Second, the estimation algorithms can be easily modified to meet specific
requirements.
Third, the (advanced) user can control estimation details generally
not available in other software packages by overriding reasonable defaults.

### 2006-08-30 paper/trunk/systemfit_summary.tex
corrected a typo in the summary
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@690 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changes according to the first major comment of reviewer A: modified title and a few sentences (package to estimate -> package for estimating; equation system -> system of equations)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@691 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary and outlook}\label{sec:Summmary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\nopagebreak
In this article, we have described some of the basic features of the
\pkg{systemfit} package for estimation of linear systems of
equations.
Many details of the estimation can be controlled by the user.
Furthermore, the package provides some statistical tests
for parameter restrictions and consistency of 3SLS estimation.
It has been tested on a variety of datasets and has produced satisfactory
for a few years.
While the \pkg{systemfit} package performs the basic fitting methods,
more sophisticated tools exist.
We hope to implement missing functionalities
in the near future.
% Some of these are discussed in the following.

@@

\subsubsection*{Non-linear estimation}
Finally, the \pkg{systemfit} package provides a function to estimate
systems of non-linear estimations.
However, the function \code{nlsystemfit} is currently under
development and the results are not yet always reliable due to
convergence difficulties.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary and outlook}\label{sec:Summmary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\nopagebreak
In this article, we have described some of the basic features of the
\pkg{systemfit} package for estimating of systems of linear
equations.
Many details of the estimation can be controlled by the user.
Furthermore, the package provides some statistical tests
for parameter restrictions and consistency of 3SLS estimation.
It has been tested on a variety of datasets and has produced satisfactory
for a few years.
While the \pkg{systemfit} package performs the basic fitting methods,
more sophisticated tools exist.
We hope to implement missing functionalities
in the near future.
% Some of these are discussed in the following.

@@

\subsubsection*{Non-linear estimation}
Finally, the \pkg{systemfit} package provides a function to estimate
systems of non-linear equations.
However, the function \code{nlsystemfit} is currently under
development and the results are not yet always reliable due to
convergence difficulties.
